---
title: "Antimicrobial-resistant Enterobacterales association with zip code: 2014-2021"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, error = FALSE, eval = TRUE, tidy = TRUE, cache = FALSE)
```

```{r libraries}
packages <- function(x){
  for(i in x){
     if(!require(i, character.only = TRUE)){
      install.packages(i, dependencies = TRUE)
       library(i, character.only = TRUE)
    }
  }
}
packages(c("readr", "readxl", "haven", "tidyverse" , "data.table" , "naniar", "rockchalk", "lubridate", "kableExtra", "DT", "formatR", "zoo", "binom", "plotly", "janitor", "sjlabelled", "Hmisc", "rms", "aod", "sandwich", "lmtest", "tableone", "broom", "gmodels", "arsenal", "investr", "lessR", "RColorBrewer"))
```

# **Raw data files**

```{r raw data files}
options(scipen=999)
setwd("C:/Users/henderh/Desktop/Projects/HAMR/HAMR_zip_Rproj")
# Read in raw data
culture_data <- readRDS("Z:/culture_data_clean_2000_2021_28nov2022.rds")
zip_data1 <- read_excel("Z:/Brown_dylan/VAN_DUIN_IRB18-3438_MRN_and_ZIP.xlsx")
zip_data2 <- read_excel("Z:/Susceptibilities_HIVStatus_June_2019_June29_2021/Patients_Summary_Susceptibilties_HIV.xlsx")
ses_data <- read_excel("zip_ses.xlsx")
```

```{r}
zip1 <- zip_data1 %>%
  clean_names() %>%
  dplyr::rename(pat_mrn_id = current_mrn,
                zip_code1 = zip) %>%
  select(-original_file_mrn)
zip2 <- zip_data2 %>%
  clean_names() %>%
  select(pat_mrn_id, zip) %>%
  dplyr::rename(zip_code2 = zip) %>%
  mutate(zip_code2 = substr(zip_code2, 1, 5))
zips <- zip1 %>%
  full_join(zip2) %>%
  mutate(zip_code = ifelse(!is.na(zip_code2), zip_code2, zip_code1)) %>%
  select(pat_mrn_id, zip_code) %>%
  filter(zip_code != "(null)")
rm(zip1,zip2)
```


```{r}
names(culture_data)
data1 <- culture_data %>%
  filter(culture_date >= as.Date("2014-01-01")) %>%
  select(pat_mrn_id, isolate_id, gender, dob, specimen_source_recode, race_recode, culture_date, organism_name, genus, antibiotic, class, nonsusceptible, resistant_to_class, mdr) %>%
  mutate(race = case_when(race_recode %in% c("asian", "native american or pacific islander") ~ "other",
                          is.na(race_recode) ~ "unknown",
                          TRUE ~ race_recode),
         race = factor(race, levels = c("white", "black", "other", "unknown")))
data2 <- data1 %>%
  select(pat_mrn_id, isolate_id, organism_name, culture_date) %>%
  distinct() %>%
  group_by(pat_mrn_id, organism_name) %>%
  slice(which.min(culture_date)) %>%
  ungroup
data3 <- data1 %>%
  inner_join(data2) %>%
  left_join(zips) %>%
  filter(!is.na(zip_code) & !zip_code %in% c("99999", "V5Z2A", "TW183", "J5T 2", "J2T 3", "E3C 1", "B5A 2"))
data4 <- data3 %>%
  left_join(ses_data %>% mutate(zip_code = as.character(zip_code))) %>%
  filter(!is.na(sum))
data5 <- data4 %>%
  group_by(pat_mrn_id, isolate_id, organism_name, mdr, zip_code, sum, percentile) %>%
  summarise() %>%
  ungroup() %>%
  mutate(ses_quar = ntile(sum,4),
         ses_quar = factor(ses_quar, levels = c(1,2,3,4), labels = c("high", "med-high", "med-low", "low"))) %>%
  select(-c(sum, percentile))
data6 <- data4 %>%
  mutate(age = round(as.numeric((culture_date - dob) / 365.25),1),
         age_50 = ifelse(age < 50, FALSE, TRUE),
         culture_year = year(culture_date),
         specimen_source = factor(specimen_source_recode, levels = c("urine", "blood", "resp", "other"))) %>%
  left_join(data5) %>%
  select(-specimen_source_recode)
data7 <- data6 %>%
  select(pat_mrn_id, isolate_id, gender, age, race, culture_year, specimen_source, mdr, ses_quar) %>%
  distinct()
data7 %>%
  tabyl(ses_quar, mdr) %>%
  adorn_totals("row") %>%
  adorn_totals("col") %>%
  adorn_percentages() %>%
  adorn_pct_formatting() %>%
  adorn_ns() %>%
  adorn_title("top", "MDR by SES quartile")
n_distinct(data7$pat_mrn_id)
# 76,118 patients
```

# **Analysis**

### **Descriptive statistics**

```{r descriptive stats}
data <- data7 %>%
  select(isolate_id, age, gender, race, specimen_source, ses_quar) %>%
  distinct()
vars <- c("age", "gender", "race", "specimen_source")
factorVars <- c("gender", "race", "specimen_source")
tableOne <- CreateTableOne(vars = vars, data = data, strata = "ses_quar", factorVars = factorVars)
tbl <- (print(tableOne, nonnormal = TRUE, quote = FALSE, noSpaces = TRUE, printToggle = FALSE, test = FALSE))
tbl
```

### **Five-year moving average prevalence of MDR**

```{r yearly moving average mdr}
a <- data7 %>%
  group_by(mdr, culture_year, ses_quar) %>%
  summarise(n = n()) %>%
  group_by(culture_year, ses_quar) %>%
  filter(mdr) %>%
  group_by(ses_quar) %>%
  arrange(culture_year, .by_group = TRUE) %>%
  mutate(num = rollmean(n, k = 5, fill = NA)) %>%
  ungroup() %>%
  select(-mdr)
b <- data7 %>%
  group_by(culture_year, ses_quar) %>%
  summarise(n = n()) %>%
  group_by(ses_quar) %>%
  arrange(culture_year, .by_group = TRUE) %>%
  mutate(den = rollmean(n, k = 5, fill = NA)) %>%
  ungroup() 
tb_mdr_5yr <- a %>%
  mutate(den = b$den,
         mn = num / den * 100,
         mn = round(mn, 2),
         class = "Multidrug-resistant")
```

#### **Chart: Five-year moving average prevalence of MDR**

```{r}
plt <- ggplot(tb_mdr_5yr, aes(x = culture_year, y = mn, colour = ses_quar)) +
  geom_point(aes(group = ses_quar), size = 1.7) +
  geom_smooth(aes(group = ses_quar), method = "loess", se = FALSE, size = 0.75) +
  theme(panel.border = element_blank(),  
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(color = "grey"),
  text = element_text(size = 12), 
  legend.position = "top", 
  legend.key = element_blank()) +
  ylim(0, 25) +
  xlim(2016, 2019) +
  xlab("Calendar year") +
  ylab("Percent multidrug-resistant")
ggplotly(plt)
```

### **Yearly average prevalence of MDR**

```{r yearly average mdr}
tb_mdr <- data7 %>%
  group_by(isolate_id, culture_year, mdr, ses_quar) %>%
  summarise() %>%
  group_by(mdr, culture_year, ses_quar) %>%
  summarise(n = n()) %>%
  group_by(culture_year, ses_quar) %>%
  mutate(total = sum(n)) %>%
  filter(mdr) %>%
  mutate(mn = round(n/total * 100, 1),
         class = "Multidrug-resistant") %>%
  ungroup() %>%
  select(-mdr)
```

#### **Chart: Yearly average prevalence of MDR**

```{r}
plt <- ggplot(tb_mdr, aes(x = culture_year, y = mn, colour = ses_quar)) +
  geom_point(size = 1.3, alpha = .5) +
  geom_smooth(aes(group = ses_quar), method = "loess", span = 1, se = FALSE, size = 0.75) +
  theme(panel.border = element_blank(),  
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(color = "grey"),
  text = element_text(size = 12), 
  legend.position = "top", 
  legend.key = element_blank()) +
  scale_y_continuous("Percent multidrug-resistant", breaks = seq(0, 30, 5), limits = c(0, 30)) +
  xlim(2014, 2021) +
  xlab("Calendar year") +
  ylab("Percent multidrug-resistant") +
  scale_color_manual(name = "",
                        breaks = c("high", "med-high", "med-low", "low"),
                        labels = c("High", "Medium high", "Medium low", "Low"),
                     values = c("darkred", "darkblue", "darkgreen", "darkgrey"))
plt
# ggsave(filename = "figure2_new.tiff", plot = plt, device="tiff", dpi=600, width = 7, height = 4)
```

### **Prevalence of resistance to antibiotic classes, by SES quartile**

```{r resistance in specific classes}
# class_factor <- c(aminoglycoside = "Aminoglycosides", cephalosporin_3rd_gen = "Cephalosporins 3rd generation", cephalosporin_2nd_gen = "Cephalosporins 2nd generation", penicillins = "Penicillins", penicillins_bli = "Penicillin + beta-lactamase inhibitor", sulfonamide = "Sulfonamides", quinolone = "Quinolones", mdr = "Multidrug-resistant")
overall_mdr <- data6 %>%
  group_by(isolate_id, mdr, ses_quar) %>%
  summarise() %>%
  group_by(mdr, ses_quar) %>%
  summarise(subgroup = n()) %>%
  group_by(ses_quar) %>%
  mutate(group = sum(subgroup),
         pct_nonsusc = (subgroup / group) * 100, 
         pct_nonsusc = round(pct_nonsusc, 1),
         class = "mdr",
         resistant_to_class = TRUE) %>%
  filter(mdr)
classes <- data6 %>%
  filter(class %in% c("aminoglycoside", "cephalosporin_2nd_gen", "cephalosporin_3rd_gen", "penicillins", "penicillins_bli", "sulfonamide", "quinolone")) %>%
  group_by(isolate_id, class, resistant_to_class, ses_quar) %>%
  summarise() %>%
  group_by(class, resistant_to_class, ses_quar) %>%
  summarise(subgroup = n()) %>%
  group_by(class, ses_quar) %>%
  mutate(group = sum(subgroup),
         pct_nonsusc = (subgroup / group) * 100, 
         pct_nonsusc = round(pct_nonsusc, 1)) %>%
  filter(resistant_to_class) %>%
  bind_rows(overall_mdr %>% select(-mdr)) %>%
  ungroup %>%
  mutate(class = factor(class, levels = c("aminoglycoside", "carbapenem", "cephalosporin_2nd_gen", "cephalosporin_3rd_gen", "penicillins", "penicillins_bli", "quinolone", "sulfonamide", "mdr")))
  # class = recode(class, !!!class_factor)) 
CIs <- binom.confint(x = classes$subgroup, n = classes$group, methods = "wilson")
classes$lower <- round(CIs$lower * 100, 1)
classes$upper <- round(CIs$upper * 100, 1)
```

#### **Chart: Percent resistant by antibiotic class and SES quartile, 2014-2021**

```{r fig.width=7, fig.height=6}
names <- c("Aminoglycosides", "Cephalosporins 2nd generation", "Cephalosporins 3rd generation",  "Penicillins", "Penicillin + beta-lactamase inhibitor", "Quinolones", "Sulfonamides", "Multdrug-resistant")
plt <- ggplot(data = classes, aes(x = class, y = pct_nonsusc, ymin = lower, ymax = upper, colour = ses_quar)) + 
  geom_point(position = position_dodge(width = 0.6)) +
  geom_errorbar(width = .3, position = position_dodge(width = 0.6)) +
  theme(panel.border = element_blank(),  
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "grey"),
  plot.caption = element_text(hjust = 0),
  text = element_text(size = 12), 
  legend.position = "top", 
  legend.key = element_blank()) +
  scale_y_continuous("Percent nonsusceptible", breaks = seq(0, 70, 10), limits = c(0, 75)) + 
  scale_x_discrete("", limits = rev,
                   breaks = c("aminoglycoside", "cephalosporin_2nd_gen", "cephalosporin_3rd_gen", "penicillins", "penicillins_bli", "quinolone", "sulfonamide", "mdr"),
                   labels = names) +
  scale_color_manual(name = "",
                        breaks = c("high", "med-high", "med-low", "low"),
                        labels = c("High", "Medium high", "Medium low", "Low"),
                     values = c("darkred", "darkblue", "darkgreen", "darkgrey")) +
  coord_flip()
plt
# ggsave(filename = "figure1_new.tiff", plot = plt, device="tiff", dpi=600, width = 7, height = 4)
```

### **Models - log-binomial**

```{r}
names(data7)

func_pr <- function(fmla) {
  model <- glm(fmla, data = data7, family = poisson(link = "log"))
  a <- (coeftest(model, vcov = sandwich))
  b <- cbind(a[,1], (confint(model)))
}

fmla <- mdr ~ ses_quar

mod_pr <- func_pr(fmla)

as.data.frame(mod_pr) %>%
  rownames_to_column("feature") %>%
  clean_names() %>%
  mutate(est = exp(v1),
         lower_cl = exp(x2_5_percent),
         upper_cl = exp(x97_5_percent)) %>%
  mutate_if(is.numeric, round, 2) %>%
  select(feature, est, lower_cl, upper_cl) %>%
  filter(feature != "(Intercept)")
```

